\large{\bf Looking forwards and backwards: dynamics and genealogies of locally regulated populations }

### 2023-12-10 forwards-and-backwards.tex
most of Mattias comments

\documentclass[EJP]{ejpecp} % replace ECP by EJP if needed.

@@

\newcommand{\DG}{\mathcal{B}}  % generator of the dispersal process
\newcommand{\DD}{\mathcal{D}}  % the second-order part of the generator of dispersal
\newcommand{\meanq}{\vec b}    % mean of dispersal, times theta
\newcommand{\covq}{\mathbf{C}}     % covariance matrix of dispersal, times theta
\newcommand{\kernel}{\rho}  % interaction kernels
\newcommand{\smooth}[1]{\kernel_{#1} \! * \!}  % convolution by the interaction kernel
\newcommand{\wavespeed}{\mathfrak{c}}    % speed (vector) of a wave
\newcommand{\Lgen}{\mathcal{L}}    % generator of a lineage
\newcommand{\Pgen}{\mathcal{P}}    % generator of the population process
\newcommand{\lp}{\xi}              % process with levels
\newcommand{\labelspace}{\mathcal{I}} % space of labels
\newcommand{\concat}{\oplus}   % concatenation of labels
\newcommand{\measures}{\mathcal{M}_F(\IR^d)} % finite measures on Rd
\newcommand{\cmeasures}{\mathcal{M}_F(\overline{\IR}^d)} % finite measures on compactified Rd
\newcommand{\lpmeasures}{\mathcal{M}(\overline{\IR^d} \times [0,\infty))} % locally finite measures on (compactified space) x level
\newcommand{\flpmeasures}{\mathcal{M}(\IR^d \times [0,\infty))} % locally finite measures on space x level

@@

\begin{remark}
\label{ancestral lineages: first guess} \revpoint{1}{7}
An equivalent way to describe the model would be to say that
when the state of the population is $\eta$,
an individual at $x$ gives birth at rate
$$
    \theta \gamma(x, \eta)
    \int r(y, \eta) q(x, dy) ,
$$
and that offspring disperse according to the kernel
$$
    q_\theta^\mathfrak{m}(x,\eta,dy)
    :=
    \frac{
        r(y, \eta) q_\theta(x, dy)
    }{
        \int r(z, \eta) q_\theta(x, dz)
    } .
$$
Clearly, the random walk driven by this dispersal kernel
is biased towards regions of higher establishment probability.
For comparison with future results,
it is interesting to write down the limiting generator:
\begin{align}
	\label{weighted dispersal}
    \lim_{\theta \to \infty}
    \theta \int (f(y) - f(x)) q_\theta^\mathfrak{m}(x, \eta, dy)
    &=
    \frac{
        \DG\left[ f(\cdot) r(\cdot, \eta) \right](x)
        - 
        f(x) \DG\left[ r(\cdot, \eta) \right](x)
    }{
        r(x, \eta^{N})
    } .
\end{align}
In the simplest case of unbiased isotropic dispersal
(i.e.,~$\meanq = 0$ and $\covq = I$), $\DG = \Delta$,
and so~(\ref{weighted dispersal}) is equal to
\begin{align*}
    \Delta f(x) + 2 \grad f(x) \cdot \grad \log r(\cdot, \smooth{r} \eta(\cdot))(x) .
\end{align*}
One might guess that the spatial motion described by following  
the ancestral lineage of an individual 
back through time
would be described (in the limit) by the adjoint of this generator.
However, we will see in Section~\ref{sec:ancestral lineages}
that this is not in fact the case.
\end{remark}

@@

\begin{proposition}
    \label{prop:nonlocal_to_local}
Let $\rho^\epsilon_F(x)=\rho_F\big(x/\epsilon)/\epsilon^d$.
Assume $\varphi_0\in L^2(\IR^d)$ is a positive, uniformly Lipschitz, and uniformly bounded function. 
Suppose that $\varphi^\epsilon\in L^2(\IR^d)$ is a weak solution to the equation
\begin{equation}
\label{nonlocalPDEv1} 
%\begin{cases}
\partial_t \varphi^\epsilon = \DG^* \varphi^\epsilon + 
\varphi^\epsilon F(\rho_F^\epsilon*\varphi^\epsilon),  
	%& 
\qquad x \in \mathbb{R}^d,\, t >0, %\\ 
%	\varphi^\epsilon(0,x) = \varphi_0(x) & x \in \mathbb{R}^d;
%\end{cases}
\end{equation}
with initial condition $\varphi_0(\cdot)$,
and that $\varphi$ is a weak solution to the equation
 \begin{equation}
\label{localPDE} 
%\begin{cases}
\partial_t \varphi = \DG^* \varphi + \varphi F(\varphi),  
%	& 
\qquad x \in \mathbb{R}^d,\, t >0, %\\ 
%	\varphi(0,x) = \varphi_0(x) & x \in \mathbb{R}^d.
%\end{cases}
\end{equation}
also with initial condition $\varphi_0(\cdot)$.
%Assume $\varphi_0\in L^2(\IR)$ is a positive, Lipschitz, and uniformly bounded function; 
Suppose further that $F$ is a Lipschitz function which is bounded above, and that
$\meanq(x)$ and $\covq(x)$, the drift and 
covariance matrix of $\DG$, satisfy the conditions of 
Assumptions~\ref{def:model_setup}.
Then, for all $T>0$ there exists a constant $K=K(T, \Vert \varphi_0 \Vert_\infty) < \infty$
and a function $\delta(\epsilon)$ (dependent on $\rho_F$) with $\delta(\epsilon)\to 0$ 
as $\epsilon\to 0$,
such that, for all $0 \leq t \leq T$, and $\epsilon$ small enough,
\[ 
\Vert \varphi_t(\cdot) - \varphi_t^\epsilon(\cdot) \Vert_\infty\leq K\delta(\epsilon). 
\]
In particular, as $\epsilon\to 0$, we have that $\varphi^\epsilon$ converges 
uniformly in compact intervals of time to $\varphi$.
\end{proposition}

@@

@@

Write $\tau_a$ for the birth time of the label $a$ and 
$\sigma_a = \lim_{u_0 \to \infty} \inf\{t \ge 0: U_a(t) > u_0\}$
the time the level hits $\infty$.
Suppose that the spatial locations and level of each line of descent $a$
solve, for $\tau_a \le t < \sigma_a$,
\begin{align}
    \label{eqn:limiting_construction}
    \begin{split}
X_a(t)
    &=
    X_a(\tau_a)
    + \int_{\tau_a}^{t}
        \beta(X_a(s), \eta_s) ds
    + \int_{\tau_a}^{t}
        J(X_a(s),\eta_s) dW_a(s)
    \\
U_a(t)
    &=
    U_a(\tau_a)
    + \int_{\tau_a}^{t}
    \bigg(
        \alpha \gamma(X_a(s),\eta_s)
        r(X_a(s), \eta_s) U_a(s)^2
\\ &\qquad \qquad \qquad {}   
        -
        \big\{
            \gamma(X_a(s),\eta_s) \DG r(X_a(s),\eta_s)
            + F(X_a(s), \eta_s)
        \big\}
        U_a(s)
    \bigg)
    ds .
    \end{split}
\end{align}
Each point in each $R_a$ denotes a potential birth time for $a$:
if the $j^\text{th}$ point in $R_a$ is $(\tau, v)$, with 
$\tau_a \le \tau < \sigma_a$,
then a new line of descent with label $a \concat j$ is produced,
with birth time $\tau_{a \concat j} = \tau$,
    location $X_{a \concat j}(\tau) = X_a(\tau)$, and level
$$
    U_{a \concat j}(\tau) = U_a(\tau)
    + \frac{v}{ 2 \alpha \gamma(X_a(\tau), \eta_\tau) r(X_a(\tau), \eta_\tau) } ,
$$
if this is finite.
For any solution to the equations above, \revpoint{1}{24}
the processes defined by
\begin{align*}
    \eta_t
    =
    \lim_{u_0 \to \infty} \frac{1}{u_0}
        \sum_{a : \tau_a \le t < \sigma_a \; ; \; U_a(t) < u_0} \delta_{X_a(t)}
    \qquad \text{and} \qquad
    \lp_t = \sum_{a : \tau_a \le t < \sigma_a} \delta_{(X_a(t), U_a(t))} 
\end{align*}
are solutions to the martingale problems of
Theorems~\ref{thm:nonlocal_convergence} and~\ref{thm:lookdown_convergence},
respectively.
\end{proposition}

@@

@@

\begin{proof}
    First note that by Lemma~\ref{lem:eta_f_bound},
    $\IE[\langle 1, \eta^N_t \rangle] \le \IE[\langle 1, \eta^N_0 \rangle] e^{Ct}$
    for some $C$ (independent of $N$).
    Now, let $M^{N*}_t(f) = \sup_{0 \le s \le t} M^N_t(f)$,
    and as before let $\langle M^N(f) \rangle_t$ be the angle bracket process of $M^N_t(f)$.
    The Burkholder-Davis-Gundy inequality says that there is a $K$ for which
    $\IE\left[ M^{N*}_t(1) \right] \le K \IE[\sqrt{[M^N(1)]_t}]$,
    where $[M^N(1)]_t$ is the quadratic variation of $M^N(1)$.
    Furthermore, as discussed by \citet{hernandez/jacka:2022},
    % for $q \le 1$ there exist universal constants $c_q$ such that
    % for any local martingale with $M_0=0$, $\IE[[M]^{q}_T] \le c_q \IE[\langle M \rangle^q_T]$
    the expectation of the quadratic variation of a local martingale
    is bounded by a (universal) constant multiple of the expectation of its angle bracket process
    (\cite{barlow/jacka/yor:1986} Item (4.b'), Table 4.1, p. 162).
    Now, since $\sqrt{x} \le 1 + x$,
    in the notation of Lemma~\ref{lem:eta_f_bound}, there is a $C'$ such that
    \begin{align*}
        \IE\left[ M^{N*}_t(1) \right]
        &\le
        C'\left( 1 + \IE\left[ \left\langle M^N(1) \right\rangle_t \right] \right)
        \\ &=
        C'\left( 1 + \frac{\theta}{N} \IE\left[
            \int_0^t
            %\int_{\IR^d} 
	    \Big\langle\left\{
                \gamma(x, \eta^N_s)
                \int_{\IR^d} r(y, \eta^N_s) q_\theta(x, dy)
                + \mu_\theta(x, \eta^N_s)
            \right\}, \eta_s^N(dx)\Big\rangle
            ds
            \right] \right)
        \\ &=
        C'\Big( 1 + \IE\Big[
            \int_0^t
            %\int_{\IR^d} 
	    \Big\langle\Big\{
                \frac{2\theta}{N} \gamma(x, \eta^N_s) r(x, \eta^N_s)
        \\ &\qquad\qquad\qquad\qquad \qquad {}
		+ \frac{\gamma(x,\eta_s^N)}{N} 
		B^\theta_1(x, \eta^N_s)-\frac{1}{N}F(x,\eta^N_s)
            \Big\}, \eta_s^N(dx) \Big\rangle
            ds
            \Big] \Big) .
    \end{align*}

@@

Furthermore, let $\Pgen^\infty$ be the generator given by
\begin{equation}
    \label{eq: Limit Generator Definition}
\begin{aligned}
\Pgen^{\infty} G_f(\eta)
    &:= G'(\langle f, \eta \rangle)
                   \big\langle
                        \gamma(x, \eta)
                            \mathcal{B}\left(
                            f(\cdot) r(\cdot, \eta)
                            \right)(x)
                    +
                    f(x) F(x, \eta),
                    \eta(dx)
                    \big\rangle \\
               &\qquad {}
               + \alpha G''(\langle f, \eta \rangle)
                  \big\langle
                    \gamma\left( x, \eta \right)
                    r\left(x,\eta \right)
                    %+\mu\left(x,\eta \right)
                    f^2(x),
                    \eta (dx)
                    \big\rangle  .
\end{aligned}    
\end{equation}
A process $(\eta^{\infty}_t)_{t \geq 0}$ 
satisifes the martingale characterization of equations~\eqref{eqn:limiting_mgale_problem}
and~\eqref{eqn:limiting_mgale_variation}
if and only if it is a solution 
to the $(\Pgen^{\infty}, \eta^\infty_0)$-martingale problem, i.e.,
if for all such test functions 
$$
    M_t :=
    G_f(\eta^{\infty}_t) - G_f(\eta^\infty_0)
    - \int_{0}^{t}\Pgen^{\infty}G_f(\eta^{\infty}_s)ds
$$
is a martingale (with respect to the natural $\sigma$-field). 
\end{lemma}

@@

\begin{lemma}
\label{no mass at infty}
Under the assumptions of Theorem~\ref{thm:nonlocal_convergence},
let $(\eta_t)_{t\geq 0}$ be a limit point of $\{(\eta^N_t)_{t\geq 0}\}_{N\geq 1}$. 
For any $\delta>0$, 
\[
\IP\big[\eta_t\big(\{\|x\|>R\}\big)>\delta\big]\to 0\qquad\mbox{ as }R\to\infty.
\]
\end{lemma}

\begin{proof}[Sketch]
Take $f_0(x)$ as in the statement of Theorem~\ref{thm:nonlocal_convergence},
i.e., $f_0$ is nonnegative, grows to infinity as $x \to \infty$,
has uniformly bounded first and second derivatives,
and has $\langle f_0, \eta^N_0\rangle$ uniformly bounded in $N$.
We take a sequence of test functions $f_n$ that increase to the function $f_0$
and having uniformly bounded first and second derivatives,
so that there is a (single) $C$
from Condition~\ref{gamma_B_condition} of Assumptions~\ref{def:model_setup}
such that $\gamma(x,\eta) B_{f_n}^\theta(x,\eta) \le C(1 + f_n(x))$
for all $x$, $\eta$, and $f_n$.
Then, just as we arrived at equation~\eqref{bound on intfdeta},
\begin{align*}
    \IE[\langle f_n(x), \eta^N_t(dx) \rangle]
    &\le
    \IE[\langle f_n(x), \eta^N_0(dx) \rangle]
    +
    C \int_0^t \IE\Big[\Big\langle f_n(x), \eta^N_s(dx) \Big\rangle \Big] ds ,
\end{align*}
with the same constant for all $n$ and all $N$.
Gronwall's inequality then implies that 
$\IE[\langle f_n, \eta^N_t\rangle] \le C'$
for some $C'$ independent of $n$, $N$, and $t \in [0,T]$.
By first taking $N \to \infty$ and then $n \to \infty$,
we find that $\IE[\langle f_0, \eta_t(dx)\rangle] \le C'$
for $t\in [0,T]$, and since $f_0 \to \infty$ as $|x| \to \infty$,
an application of Markov's inequality
tells us that for any $\delta > 0$,
$$
    \IP\big\{ \eta_t(\{x : \|x\| > R\})>\delta \big\}
    \le \frac{\IE[\langle f_0, \eta_t\rangle]}{\inf_{\{x:\|x\|\ge R\}} f_0(x)}
    \to 0 \qquad \text{as } R\to\infty .
$$
\end{proof}

@@

    First, note that for any $\epsilon > 0$, by tightness of ${X^{(n)}_t}_{n\ge0}$ and ${Y^{(n)}_t}_{n\ge0}$
    there is a $K$ such that
    $\IP\{ X^{(n)}_t > \sqrt{K} \}$ and $\IP\{ Y^{(n)}_t > \sqrt{K} \}$
    are both less than $\epsilon/2$,
    and hence
    \begin{align*}
        \IP\{
            X^{(n)}_t Y^{(n)}_t > K
        \}
        \le
        \IP\{
            X^{(n)}_t > \sqrt{K}
        \}
        +
        \IP\{
            Y^{(n)}_t > \sqrt{K}
        \} 
        \le \epsilon.
    \end{align*}
    Therefore, ${X^{(n)}_t Y^{(n)}_t}_{n \ge 0}$ is tight.

>>>>

\RequirePackage{xcolor}
\documentclass[EJP]{ejpecp} % replace ECP by EJP if needed.

@@

\newcommand{\DG}{\mathcal{B}}  % generator of the dispersal process
\newcommand{\DD}{\mathcal{D}}  % the second-order part of the generator of dispersal
\newcommand{\meanq}{\vec b}    % mean of dispersal, times theta
\newcommand{\covq}{\mathbf{C}}     % covariance matrix of dispersal, times theta
\newcommand{\kernel}{\rho}  % interaction kernels
\newcommand{\smooth}[1]{\kernel_{#1} \! * \!}  % convolution by the interaction kernel
\newcommand{\wavespeed}{\mathfrak{c}}    % speed (vector) of a wave
\newcommand{\Lgen}{\mathcal{L}}    % generator of a lineage
\newcommand{\Pgen}{\mathcal{P}}    % generator of the population process
\newcommand{\lp}{\xi}              % process with levels
\newcommand{\labelspace}{\mathcal{I}} % space of labels
\newcommand{\concat}{\oplus}   % concatenation of labels
\newcommand{\measures}{\mathcal{M}_F(\IR^d)} % finite measures on Rd
\newcommand{\cmeasures}{\mathcal{M}_F(\overline{\IR}^d)} % finite measures on compactified Rd
\newcommand{\lpmeasures}{\mathcal{M}(\overline{\IR^d} \times [0,\infty))} % locally finite measures on (compactified space) x level
\newcommand{\flpmeasures}{\mathcal{M}(\IR^d \times [0,\infty))} % locally finite measures on space x level
\newcommand{\comment}[1]{{\color{blue} \it #1}}

@@

\begin{remark}
\label{ancestral lineages: first guess}
An equivalent way to describe the model would be to say that
when the state of the population is $\eta$,
an individual at $x$ gives birth at rate
$$
    \theta \gamma(x, \eta)
    \int r(y, \eta) q(x, dy) ,
$$
and that offspring disperse according to the kernel
$q_\theta^\mathfrak{m}$ (with superscript $\mathfrak{m}$ because it is post-\textbf{m}ortality),
defined by: \revpoint{1}{7}
$$
    q_\theta^\mathfrak{m}(x,\eta,dy)
    :=
    \frac{
        r(y, \eta) q_\theta(x, dy)
    }{
        \int r(z, \eta) q_\theta(x, dz)
    } .
$$
Clearly, the random walk driven by this dispersal kernel
is biased towards regions of higher establishment probability.
For comparison with future results,
it is interesting to write down the limiting generator:
\begin{align}
	\label{weighted dispersal}
    \lim_{\theta \to \infty}
    \theta \int (f(y) - f(x)) q_\theta^\mathfrak{m}(x, \eta, dy)
    &=
    \frac{
        \DG\left[ f(\cdot) r(\cdot, \eta) \right](x)
        - 
        f(x) \DG\left[ r(\cdot, \eta) \right](x)
    }{
        r(x, \eta)
    } .
\end{align}
In the simplest case of unbiased isotropic dispersal
(i.e.,~$\meanq = 0$ and $\covq = I$), $\DG = \Delta$,
and so~(\ref{weighted dispersal}) is equal to
\begin{align*}
    \Delta f(x) + 2 \grad f(x) \cdot \grad \log r(\cdot, \smooth{r} \eta(\cdot))(x) .
\end{align*}
One might guess that the spatial motion described by following  
the ancestral lineage of an individual 
back through time
would be described (in the limit) by the adjoint of this generator.
However, we will see in Section~\ref{sec:ancestral lineages}
that this is not in fact the case.
\end{remark}

@@

\begin{proposition}
    \label{prop:nonlocal_to_local}
Let $\rho^\epsilon_F(x)=\rho_F\big(x/\epsilon)/\epsilon^d$.
Assume $\varphi_0\in L^2(\IR^d)$ is a positive, uniformly Lipschitz, and uniformly bounded function. 
Suppose that $\varphi^\epsilon\in L^2(\IR^d)$ is a weak solution to the equation
\begin{equation}
\label{nonlocalPDEv1} 
%\begin{cases}
\partial_t \varphi^\epsilon = \DG^* \varphi^\epsilon + 
\varphi^\epsilon F(\rho_F^\epsilon*\varphi^\epsilon),  
	%& 
\qquad x \in \mathbb{R}^d,\, t >0, %\\ 
%	\varphi^\epsilon(0,x) = \varphi_0(x) & x \in \mathbb{R}^d;
%\end{cases}
\end{equation}
with initial condition $\varphi_0(\cdot)$,
and that $\varphi$ is a weak solution to the equation
 \begin{equation}
\label{localPDE} 
%\begin{cases}
\partial_t \varphi = \DG^* \varphi + \varphi F(\varphi),  
%	& 
\qquad x \in \mathbb{R}^d,\, t >0, %\\ 
%	\varphi(0,x) = \varphi_0(x) & x \in \mathbb{R}^d.
%\end{cases}
\end{equation}
also with initial condition $\varphi_0(\cdot)$.
%Assume $\varphi_0\in L^2(\IR)$ is a positive, Lipschitz, and uniformly bounded function; 
Suppose further that $F$ is a Lipschitz function which is bounded above,
and that $\meanq(x)$ and $\covq(x)$, the drift and covariance matrix of $\DG$,
satisfy the conditions of Assumptions~\ref{def:model_setup}
and are such that $\DG^* 1 = 0$ (see Definition~\ref{def:dispersal_generator}).
Then, for all $T>0$ there exists a constant $K=K(T, \Vert \varphi_0 \Vert_\infty) < \infty$
and a function $\delta(\epsilon)$ (dependent on $\rho_F$) with $\delta(\epsilon)\to 0$ 
as $\epsilon\to 0$,
such that, for all $0 \leq t \leq T$, and $\epsilon$ small enough,
\[ 
\Vert \varphi_t(\cdot) - \varphi_t^\epsilon(\cdot) \Vert_\infty\leq K\delta(\epsilon). 
\]
In particular, as $\epsilon\to 0$, we have that $\varphi^\epsilon$ converges 
uniformly in compact intervals of time to $\varphi$.
\end{proposition}

@@

\begin{remark}
    Instead of putting the fairly strong constraint that $\DG^* 1(x) = 0$ for all $x$,
    it would be enough to assume instead that $\DG^* 1(x)$ is uniformly bounded,
    so that $f \mapsto \DG^*f - f \DG^* 1$ is the generator of a conservative diffusion.
    Since then $\varphi$ solves
    $$ \partial_t \varphi = (\DG^* \varphi - \varphi \DG^* 1) + \varphi (F(\varphi) + \DG^* 1) , $$
    the proof goes through essentially unchanged,
    with only $F(\varphi(x))$ replaced with $F(\varphi(x)) + \DG^*1(x)$,
    a bounded perturbation.
\end{remark}

@@

Write $\tau_a$ for the birth time of the label $a$ and 
$\sigma_a = \lim_{u_0 \to \infty} \inf\{t \ge 0: U_a(t) > u_0\}$
the time the level hits $\infty$.
Suppose that the spatial locations and level of each line of descent $a$
solve, for $\tau_a \le t < \sigma_a$,
\begin{align}
    \label{eqn:limiting_construction}
    \begin{split}
X_a(t)
    &=
    X_a(\tau_a)
    + \int_{\tau_a}^{t}
        \beta(X_a(s), \eta_s) ds
    + \int_{\tau_a}^{t}
        J(X_a(s),\eta_s) dW_a(s)
    \\
U_a(t)
    &=
    U_a(\tau_a)
    + \int_{\tau_a}^{t}
    \bigg(
        \alpha \gamma(X_a(s),\eta_s)
        r(X_a(s), \eta_s) U_a(s)^2
\\ &\qquad \qquad \qquad {}   
        -
        \big\{
            \gamma(X_a(s),\eta_s) \DG r(X_a(s),\eta_s)
            + F(X_a(s), \eta_s)
        \big\}
        U_a(s)
    \bigg)
    ds ,
    \end{split}
\end{align}
where
\begin{align*}
    \eta_t
    :=
    \lim_{u_0 \to \infty} \frac{1}{u_0}
        \sum_{a : \tau_a \le t < \sigma_a \; ; \; U_a(t) < u_0} \delta_{X_a(t)} .
\end{align*}
Each point in each $R_a$ denotes a potential birth time for $a$:
if the $j^\text{th}$ point in $R_a$ is $(\tau, v)$, with 
$\tau_a \le \tau < \sigma_a$,
then a new line of descent with label $a \concat j$ is produced,
with birth time $\tau_{a \concat j} = \tau$,
    location $X_{a \concat j}(\tau) = X_a(\tau)$, and level
$$
    U_{a \concat j}(\tau) = U_a(\tau)
    + \frac{v}{ 2 \alpha \gamma(X_a(\tau), \eta_\tau) r(X_a(\tau), \eta_\tau) } ,
$$
if this is finite.
For any solution to the equations above, \revpoint{1}{24}
the process $\eta_t$
is a solution to the martingale problem of Theorem~\ref{thm:nonlocal_convergence},
and the process
\begin{align*}
    \lp_t = \sum_{a : \tau_a \le t < \sigma_a} \delta_{(X_a(t), U_a(t))} 
\end{align*}
is a solution to the martingale problem of
Theorem~\ref{thm:lookdown_convergence}.
\end{proposition}

@@

\comment{TODO: add note about well-definedness of $\eta_t$.}

@@

\begin{proof}
    First note that by Lemma~\ref{lem:eta_f_bound},
    $\IE[\langle 1, \eta^N_t \rangle] \le \IE[\langle 1, \eta^N_0 \rangle] e^{Ct}$
    for some $C$ (independent of $N$).
    Now, let $M^{N*}_t(f) = \sup_{0 \le s \le t} M^N_t(f)$,
    and as before let $\langle M^N(f) \rangle_t$ be the angle bracket process of $M^N_t(f)$.
    The Burkholder-Davis-Gundy inequality says that there is a $K$ for which
    $\IE\left[ M^{N*}_t(1) \right] \le K \IE[\sqrt{[M^N(1)]_t}]$,
    where $[M^N(1)]_t$ is the quadratic variation of $M^N(1)$.
    Furthermore, as discussed by \citet{hernandez/jacka:2022},
    % for $q \le 1$ there exist universal constants $c_q$ such that
    % for any local martingale with $M_0=0$, $\IE[[M]^{q}_T] \le c_q \IE[\langle M \rangle^q_T]$
    the expectation of the quadratic variation of a local martingale
    is bounded by a (universal) constant multiple of the expectation of its angle bracket process
    (\cite{barlow/jacka/yor:1986}, Item (4.b'), Table 4.1, p. 162).
    Now, since $\sqrt{x} \le 1 + x$,
    in the notation of Lemma~\ref{lem:eta_f_bound}, there is a $C'$ such that
    \begin{align*}
        \IE\left[ M^{N*}_t(1) \right]
        &\le
        C'\left( 1 + \IE\left[ \left\langle M^N(1) \right\rangle_t \right] \right)
        \\ &=
        C'\left( 1 + \frac{\theta}{N} \IE\left[
            \int_0^t
            %\int_{\IR^d} 
	    \Big\langle\left\{
                \gamma(x, \eta^N_s)
                \int_{\IR^d} r(y, \eta^N_s) q_\theta(x, dy)
                + \mu_\theta(x, \eta^N_s)
            \right\}, \eta_s^N(dx)\Big\rangle
            ds
            \right] \right)
        \\ &=
        C'\Big( 1 + \IE\Big[
            \int_0^t
            %\int_{\IR^d} 
	    \Big\langle\Big\{
                \frac{2\theta}{N} \gamma(x, \eta^N_s) r(x, \eta^N_s)
        \\ &\qquad\qquad\qquad\qquad \qquad {}
		+ \frac{\gamma(x,\eta_s^N)}{N} 
		B^\theta_1(x, \eta^N_s)-\frac{1}{N}F(x,\eta^N_s)
            \Big\}, \eta_s^N(dx) \Big\rangle
            ds
            \Big] \Big) .
    \end{align*}

@@

Furthermore, let $\Pgen^\infty$ be the generator given by
\begin{equation}
    \label{eq: Limit Generator Definition}
\begin{aligned}
\Pgen^{\infty} G_f(\eta)
    &:= G'(\langle f, \eta \rangle)
                   \big\langle
                        \gamma(x, \eta)
                            \mathcal{B}\left(
                            f(\cdot) r(\cdot, \eta)
                            \right)(x)
                    +
                    f(x) F(x, \eta),
                    \eta(dx)
                    \big\rangle \\
               &\qquad {}
               + \alpha G''(\langle f, \eta \rangle)
                  \big\langle
                    \gamma\left( x, \eta \right)
                    r\left(x,\eta \right)
                    %+\mu\left(x,\eta \right)
                    f^2(x),
                    \eta (dx)
                    \big\rangle  .
\end{aligned}    
\end{equation}
A process $(\eta^{\infty}_t)_{t \geq 0}$ 
satisfies the martingale characterization of equations~\eqref{eqn:limiting_mgale_problem}
and~\eqref{eqn:limiting_mgale_variation}
if it is a solution % "and only if" requires continuity
to the $(\Pgen^{\infty}, \eta^\infty_0)$-martingale problem, i.e.,
if for all such test functions 
$$
    M_t :=
    G_f(\eta^{\infty}_t) - G_f(\eta^\infty_0)
    - \int_{0}^{t}\Pgen^{\infty}G_f(\eta^{\infty}_s)ds
$$
is a martingale (with respect to the natural $\sigma$-field). 
\end{lemma}

The converse -- that any solution to equations~\eqref{eqn:limiting_mgale_problem}
and~\eqref{eqn:limiting_mgale_variation}
is a solution to the $(\Pgen^{\infty}, \eta^\infty_0)$-martingale problem --
requires continuity, which we expect to be true, but have not proved.

@@

\begin{lemma}
\label{no mass at infty}
Under the assumptions of Theorem~\ref{thm:nonlocal_convergence},
if $(\eta_t)_{t\geq 0}$ be a limit point of $\{(\eta^N_t)_{t\geq 0}\}_{N\geq 1}$,
then for any $\delta>0$ and $T>0$,
\[
\IP\bigg[
    \sup_{0 \le t \le T} \eta_t\big(\{\|x\|>R\}\big)>\delta
\bigg]\to 0\qquad\mbox{ as }R\to\infty.
\]
\end{lemma}

\begin{proof}[Sketch]
Take $f_0(x)$ as in the statement of Theorem~\ref{thm:nonlocal_convergence},
i.e., $f_0$ is nonnegative, grows to infinity as $x \to \infty$,
has uniformly bounded first and second derivatives,
and has $\langle f_0, \eta^N_0\rangle$ uniformly bounded in $N$.
We take a sequence of test functions $f_n$ that increase to the function $f_0$
and having uniformly bounded first and second derivatives,
so that there is a (single) $C$
from Condition~\ref{gamma_B_condition} of Assumptions~\ref{def:model_setup}
such that $\gamma(x,\eta) B_{f_n}^\theta(x,\eta) \le C(1 + f_n(x))$
for all $x$, $\eta$, and $f_n$.
Then, just as we arrived at equation~\eqref{bound on intfdeta},
\begin{align*}
    \IE\left[\left\langle f_n(x), \eta^N_t(dx) \right\rangle\right]
    &\le
    \IE\left[\left\langle f_n(x), \eta^N_0(dx) \right\rangle\right]
    +
    C \int_0^t \IE\Big[\Big\langle f_n(x), \eta^N_s(dx) \Big\rangle \Big] ds ,
\end{align*}
with the same constant for all $n$ and all $N$.
Gronwall's inequality then implies that 
$\IE[\langle f_n, \eta^N_t\rangle] \le C'$
for some $C'$ independent of $n$, $N$, and $t \in [0,T]$.
By first taking $N \to \infty$ and then $n \to \infty$,
we find that $\IE[\langle f_0, \eta_t(dx)\rangle] \le C'$ for $t\in [0,T]$.
However, this is for a single time --
we would like instead to uniformly bound
$\IE[\sup_{0 \le t \le T} \langle f_0, \eta_t(dx)\rangle]$.
This can be done in a similar but lengthier manner,
following the proof of Lemma~\ref{lem:eta_compact_containment}
and observing that bounds can be taken independent of $n$ and $N$.

Finally, since $f_0 \to \infty$ as $|x| \to \infty$,
an application of Markov's inequality
tells us that for any $\delta > 0$,
$$
    \IP\bigg\{ \sup_{0 \le t \le T} \eta_t(\{x : \|x\| > R\})>\delta \bigg\}
    \le \frac{\IE[\sup_{0 \le t \le T} \langle f_0, \eta_t\rangle]}{\delta \inf_{\{x:\|x\|\ge R\}} f_0(x)}
    \to 0 \qquad \text{as } R\to\infty .
$$
\end{proof}

@@

    First, note that for any $\epsilon > 0$, by tightness of $(X^{(n)}_t)_{n\ge0}$ and $(Y^{(n)}_t)_{n\ge0}$
    there is a $K$ such that
    $\IP\{ X^{(n)}_t > \sqrt{K} \}$ and $\IP\{ Y^{(n)}_t > \sqrt{K} \}$
    are both less than $\epsilon/2$,
    and hence
    \begin{align*}
        \IP\{
            X^{(n)}_t Y^{(n)}_t > K
        \}
        \le
        \IP\{
            X^{(n)}_t > \sqrt{K}
        \}
        +
        \IP\{
            Y^{(n)}_t > \sqrt{K}
        \} 
        \le \epsilon.
    \end{align*}
    Therefore, $(X^{(n)}_t Y^{(n)}_t)_{n \ge 0}$ is tight.
